### 2018-05-30 5_eval.tex
Complete everything. Left only polishing.

Figure~\ref{dep:events} illustrates the dependency of overall number of events in all unrolled event-flow graphs on the unrolling bound $k$ for the \texttt{Dekker} test.
%The left-hand side picture presents the dependency graph for \porthos[2], 
Both charts are limited above by $420$ events.
\porthos[2] (left-hand chart) reaches this limit with the bound $k_2=32$, and \porthos[1] (right-hand chart) reaches it with $k_1=5$.
However, taking into account the fact that initially both non-unrolled graphs had $20$ events, \porthos[2] at the step $32$ (the longest execution) obtained same number of events  as \porthos[1] at the step $20*5=100$, which shows that it has lost approx. $70$ states.

\begin{figure}[t]
\begin{minipage}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-events-new.png}
\end{minipage}
\hfill
%
\begin{minipage}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-events-old.png}
\end{minipage}
%
\caption{Dependency of number of events in the unrolled program on the unrolling bound}
\label{dep:events}
\end{figure}

@@

As an example of the C program liable for the reachability and portability analysis, consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.

For time benchmarking we ran the tool $5$ times and computed the median of the encoding time.
Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
The time was measured by the tool itself via the native Java method \texttt{System.currentTimeMillis}.

@@

%\textbf{TODOOOOOOOOOOOOOOOOOOOOO!!!! Re-compute tests with new counting of encoding}
The sample output of \porthos[2] in the state reachability analysis mode is presented in Appendix~\ref{apx:output}.
The output includes detailed time measurements for each stage of the analysis, and optionally statistical information about the unrolled program (number of events, transitions, etc.)

\begin{figure}[t]
\begin{minipage}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-time-new.png}
\end{minipage}
\hfill
%
\begin{minipage}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-time-old.png}
\end{minipage}
%
\caption{Dependency of program domain encoding time (in seconds) on the unrolling bound}
\label{dep:time}
\end{figure}

%For \porthos[1] we specify the unrolling bound $k_1 = 2$, and for \porthos[2] the bound is $k_2 = 21$ ().
As the unrolling algorithm has been changed, the number of events after unrolling differs considerably.
Therefore, the correct performance comparison with \porthos[1] is not manageable as the performance of the tools depends directly on number of events.
For example, with the bound $k_1=2$ \porthos[1] produces $59$ events (among them, $51$ shared-memory events) with the program domain encoding time 4.608 sec,
and with the bound $k_1=3$ \porthos[1] produces $95$ events (among them, $107$ shared-memory events) with the program domain encoding time 21.475 sec;
whereas \porthos[2] with the bound $k_2=17$ produces $57$ ($31$) events with the program domain encoding time 1.734, and with the bound $k_2=27$ it produces $193$ events (among them, $107$ shared-memory events, which equal to the number of shared-memory events produced by \porthos[1] with the bound $k_1=3$) with the time 32.673 sec.
Therefore, the new unrolling scheme has made the analysis complete but at the same time it reduced the performance considerably.
%The last result shows that with the bounds $k_1=3$ and $k_2=27$ both tools have executed same number of loops  almost equivalent as they
%However, the program domain encoding and memory model encoding schemes have not been changed considerably comparing to \porthos[1], therefore the execution time 
%Although, the performance depends directly on number of events,

@@

Figure~\ref{dep:time} illustrates the dependency of overall time spent for the program domain encoding on the unrolling bound.
The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.

Note that the time spent by \porthos[2] for the interpretation and unrolling stages remains to be negligible comparing to the full encoding time.
Therefore, we conclude that the new architecture implies no performance overhead comparing to the previous version of the tool.
The time spent by the SMT-solver while considering the SMT-formula encoded by \porthos[2] was 0.039 sec, and in case of \porthos[1] the time was 1.291 sec.
The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.

@@

For evaluating the tool working in the portability analysis mode, we used the same file \texttt{Dekker.c} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
%!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 65.693 sec is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).

>>>>

% ---------------------------

\begin{figure}[t]
\begin{subfigure}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/new/new-e.png}
\caption{The graph unrolled by \porthos[2]}
\label{dep:events:new}
\end{subfigure}
\hfill
%
\begin{subfigure}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/new/old-e.png}
\caption{The graph unrolled by \porthos[1]}
\label{dep:events:old}
\end{subfigure}
%
\caption{Dependency of number of events in the unrolled Dekker's algorithm on the unrolling bound}
\label{dep:events}
\end{figure}

As another example of program liable for analysis consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
For performing tests, \porthos[2] was slightly modified in the branch \textit{ay/test-timing}%
\footnote{The modifications were pushed into the \porthos[1] repository \url{github.com/hernanponcedeleon/Dat3M}} %
%
for printing detailed timing information and statistics about number of events (commit f8a6c97f) and writing this information in strict \textit{json} format so that it can be easily processed later by a script-analyser (commit 48ba7161).
The version of \porthos[2] used for the tests is actual up to commit 01dde6b1 of the \porthos[2] repository%
\footnote{\porthos[2] repository: \url{https://github.com/ajuszkowski/PorthosC}}. %
%
Both \porthos[1] and \porthos[2] are run in the state reachability mode for \texttt{TSO} memory model (see sample output of \porthos[2] in Appendix~\ref{apx:output}).

Figure~\ref{dep:events} illustrates the dependency of overall number of events in all unrolled event-flow graphs of the thread \texttt{t0} on the unrolling bound $k$ for the \texttt{Dekker} test (the left-hand side picture illustrates the case for \porthos[2], and the right-hand side picture illustrates the case for \porthos{}).
The solid lines denote dependency of number of unrolled events on the unrolling bound, when the dashed lines denote the number of events in non-unrolled graphs.
%Both charts have limits above $420$ events.
%\porthos[2] (left-hand chart) reaches this limit with the bound $k_2=32$, and \porthos[1] (right-hand chart) reaches it with $k_1=5$.
%However, taking into account the fact that initially both non-unrolled graphs had $20$ events, \porthos[2] at the step $32$ (the longest execution) obtained same number of events  as \porthos[1] at the step $20*5=100$, which shows that it has lost approx. $70$ states.
The blue line depicts the dynamics of the overall number of events in the graph (including memory events, nop events, etc.).
Initially, the thread \texttt{t0} has 22 events.
As in the unrolling scheme of \porthos[1], each loop is executed exactly $k_1$ times, therefore if $k_1=6$, the loop has been executed at least 6 times by producing around $300$ events.
Nonetheless, in the new unrolling scheme used by \porthos[2], the same number of events is produced by the unrolling bound $k_2=34$, which contains approximately $1.5$ executions of the $22$-event loop.
Therefore, according to rough estimates the new encoding scheme produces $6/1.5=4$ times more events than the older one.

The same dynamics can be seen with shared memory events (yellow lines).
However, the new algorithm produces much more local memory events (green line on the left-hand side picture), which grow exponentially with the unrolling bound, when the number of local memory events produced by the old algorithm (green line in the right-hand side picture) grows relatively slowly.
This can be explained by the need to copy all shared variables to local ones when performing computation over them.
On the other hand, the right-hand side picture reveals exponential-like growth of the number of \texttt{nop}- (or \texttt{skip}-) events, that are necessary for implementing the control-flow as an AST, however constitute a redundant part of the model (in the new unrolling scheme, the number of \texttt{nop}-events remains to be relatively constant).

As the new unrolled graph has all possible executions unlike the older one, by default \porthos[2] uses the new unrolling scheme.
However, in some applications the older unrolling mode may be useful (for example, where the user can use other instruments to prove the limitations on the number of executions of a loop), thus the user should be able to change configure the unrolling algorithm (support for the older algorithm is left to a future work).

@@

Performance evaluation has also been performed on the example of the \texttt{Dekker} test.
%As an example of the C program liable for the reachability and portability analysis, consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.
%For testing \porthos[1], we used the same file in old \porthos{} input language (\texttt{dekker.pts}), which was used in evaluation tests in the original paper~\cite{Porthos17a}.
For time benchmarking we ran the tools $5$ times and computed the median of the encoding time.
Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
The time was measured by the tool itself via the native Java method \texttt{System.currentTimeMillis}.

@@

The sample output of \porthos[2] in the state reachability analysis mode is presented in Appendix~\ref{apx:output}.
The output includes detailed time measurements for each stage of the analysis, and optionally statistical information about the unrolled program (number of events, transitions, etc.)

\begin{figure}[t]
\begin{subfigure}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/new/new-t.png}
\caption{Encoding of the graph unrolled by \porthos[2]}
\label{dep:time:new}
\end{subfigure}
\hfill
%
\begin{subfigure}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/new/old-t.png}
\caption{Encoding of the graph unrolled by \porthos[1]}
\label{dep:time:old}
\end{subfigure}
%
\caption{Dependency of program domain encoding time (in seconds) on the unrolling bound}
\label{dep:time}
\end{figure}

As the unrolling algorithm has been changed, the number of events after unrolling differs considerably, therefore the correct performance comparison with \porthos[1] is not manageable as the performance of the tools depends directly on number of events.

%For example, with the bound $k_1=2$ \porthos[1] produces $59$ events (among them, $51$ shared-memory events) with the program domain encoding time 4.608 sec,
%and with the bound $k_1=3$ \porthos[1] produces $95$ events (among them, $107$ shared-memory events) with the program domain encoding time 21.475 sec;
%whereas \porthos[2] with the bound $k_2=17$ produces $57$ ($31$) events with the program domain encoding time 1.734, and with the bound $k_2=27$ it produces $193$ events (among them, $107$ shared-memory events, which equal to the number of shared-memory events produced by \porthos[1] with the bound $k_1=3$) with the time 32.673 sec.
%Therefore, the new unrolling scheme has made the analysis complete but at the same time it reduced the performance considerably.

%The last result shows that with the bounds $k_1=3$ and $k_2=27$ both tools have executed same number of loops  almost equivalent as they
%However, the program domain encoding and memory model encoding schemes have not been changed considerably comparing to \porthos[1], therefore the execution time 
%Although, the performance depends directly on number of events,

@@

Figure~\ref{dep:time} illustrates the dependency of the time spent for the program domain encoding (for the process \texttt{t0}) on the unrolling bound.
Since the program domain encoding time is heavily dependent on the number of events (specifically, shared and local memory events as the encoding function contains several multiple nested loops over memory events), the dependency graph follows the results presented in Figure~\ref{dep:events} and resembles exponential relationship between encoding time and unrolling bound.

%The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.
An example with detailed execution time information for \porthos[2] can be found in Appendix~\ref{apx:output} (part "\texttt{timers}" of the output).
Note that the time spent for the interpretation (0.222 sec) and unrolling (0.088 sec) stages remains negligible comparing to the full execution time (40.679 sec).
Therefore, we conclude that the new architecture implies no performance overhead comparing to the previous version of the tool.
%The time spent by the SMT-solver while considering the SMT-formula encoded by \porthos[2] is 0.397 sec, and in case of \porthos[1] the time was 1.291 sec.
%The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
%Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.

@@

For evaluating the tool working in the portability analysis mode, we used the same files \texttt{Dekker.c} and \texttt{Dekker.pts} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
%!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 65.693 sec is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).

### 2018-05-30 90_summary.tex
Complete everything. Left only polishing.

  \item Since \porthos[2] compiles the program AST to the \xgraph{} before encoding it into the SMT-formula, we decided to change the \textit{unrolling algorithm} from unwinding all loops $k$ times (where $k$ is the user-specified unrolling bound) to the DFS-based algorithm that explores \textit{all possible executions} that the program can produce within $k$ steps (see discussion on the unrolling in Sections~\ref{ch:impl:proc:x-unroll} and~\ref{ch:eval:show:compil}).
  Although the new unrolling algorithm produces considerably more executions than the old one and thus requires more time for analysis, \porthos[2] uses the new unrolling algorithm by default as it is complete (finds all possible executions).

@@

\item For justifying the correctness of the verification that bases on the result of the input program unrolling, one might want to be aware whether an execution in the unrolled event-flow graph has run the loop a whole number or it was interrupted on non-last event of the loop.
This can be modelled with two kinds of sink events (completed and uncompleted) of the event-flow graph as it was discussed in Sections~\ref{ch:enc:bmc:cf} and~\ref{ch:impl:proc:x-unroll}.

>>>>

  \item Since \porthos[2] compiles the program AST to the \xgraph{} before encoding it into the SMT-formula, we decided to change the \textit{unrolling algorithm} from unwinding all loops $k$ times (where $k$ is the user-specified unrolling bound) to the DFS-based algorithm that explores \textit{all possible executions} that the program can produce within $k$ steps (see discussion on the unrolling in Sections~\ref{ch:impl:proc:x-unroll} and~\ref{ch:eval:show:compil}).
  On the other hand, it is shown in Section~\ref{ch:eval:show:unrol} that the implementation of the new unrolling algorithm does not produce exponentially many dummy \texttt{nop}-events as the implementation of the old unrolling algorithm does.
  Another benefit of the new unrolling scheme is that it is much more configurable (i.e., the number of events grows gradually as the unrolling bound is being increased, see Figure~\ref{dep:events}). \\
  Although the new unrolling algorithm produces considerably more executions than the old one and thus requires more time for analysis, \porthos[2] uses the new unrolling algorithm by default as it is complete (finds all possible executions).

@@

\item For justifying the correctness of the verification that bases on the result of the input program unrolling (via the new unrolling scheme discussed in Section~\ref{ch:impl:proc:x-unroll}), one might want to be aware whether the execution in the unrolled event-flow graph has run the loop a whole number or it was interrupted on non-last event of the loop.
This can be modelled with two kinds of sink events (completed and uncompleted) of the event-flow graph as it was discussed in Sections~\ref{ch:enc:bmc:cf} and~\ref{ch:impl:proc:x-unroll}.
